\LARGE How coexistence mechanisms mediate temporal stability

### 2016-02-23 manuscript/manuscript.Rmd
adds conceptual model figure

```{r figure_1, dependson="plot-options", fig.cap="Diagram of our semi-discrete consumer-resource model. Solid lines show continuous dynamics while dashed lines show the discrete transitions.", cache=FALSE, results='hide', fig.width=4, fig.height=6, eval=FALSE}
library(png)
library(grid)
img <- readPNG("ConceptFig.png")
grid.raster(img)
```

@@

At the beginning of each season we start with initial conditions defined as $V_{t}$, $W_{t}$, and $Z_{t}$ for the dormant state, the live state, and the resource, respectively. So for each season, Eqs. 1-3 are solved given the initial conditions:
\begin{align}
  D_{i}(0) &= V_{i,t} \\
  N_{i}(0) &= W_{i,t} \\
  R(0) &= Z_{t}
\end{align}
\noindent The consumers transition between $N$ and $D$ instantaneously between years. We assume resource density does not change between years. So, at the yearly transition:
\begin{align}
  V_{i,t+1} &= [N_{i}(T^-)+D_{i}(T^-)](1-g_{t}) \\
  W_{i,t+1} &= [N_{i}(T^-)+D_{i}(T^-)]g_{t} \\
  Z_{t+1} &= R(T^-) + R(T^+)
\end{align}
\noindent where $D(T^-)$, $N(T^-)$, and $R(T^-)$ are the densities of each state at the end of the year and $g$ is a time-fluctuating activation rate that regulates how much dormant biomass is converted to growing-season live biomass each year. $R(T^+)$ is a randomly generated resource pulse from a log-normal distribution with mean $R_\mu$ and variance $R_{\sigma^2}$ (same mean and variance used to generate daily pulses). Our formulation assumes that at the end of each season all accumulated living biomass [$N(T^-)$] is converted to dormant biomass.

@@

### Model Simulations
For each coexistence mechanism (storage effect, relative nonlinearity), we ran model simulations at different levels of resource variance ($R_{\sigma^2}$), environmental cue variance ($\sigma_E$), and correlation of species responses to the environmental cue ($\rho$). 
We ran simulations for 2,000 'years' with 60 day growing seasons. 
After discarding an initial 100 years to reduce transient effects on our results, we calculated the stability of summed species biomass through time and species synchrony through time.
For storage effect simulations, species were equivalent in their ability to capture and assimilate the resource (*R*) but had different, depending on $\rho$, germination rates.
For relative nonlinearity simulations, species had equivalent germination rates ($\rho = 1$) but different resource acquisition functions (Fig. X).
We explored a limited set of parameter space where the storage effect and relative nonlinearity were operating simultaneously.

>>>>

@@

At the beginning of each season we start with initial conditions defined as $V_{t}$, $W_{t}$, and $Z_{t}$ for the dormant state, the live state, and the resource, respectively. So for each season, eqs. 1-3 are solved given the initial conditions:
\begin{align}
  D_{i}(0) &= V_{i,t} \\
  N_{i}(0) &= W_{i,t} \\
  R(0) &= Z_{t}
\end{align}
\noindent The consumers transition between $N$ and $D$ instantaneously between years. So, at the yearly transition:
\begin{align}
  V_{i,t+1} &= [N_{i}(T^-)+D_{i}(T^-)](1-g_{t}) \\
  W_{i,t+1} &= [N_{i}(T^-)+D_{i}(T^-)]g_{t} \\
  Z_{t+1} &= R(T^-) + R(T^+)
\end{align}
\noindent where $D(T^-)$, $N(T^-)$, and $R(T^-)$ are the densities of each state at the end of the year and $g$ is a time-fluctuating activation rate that regulates how much dormant biomass is converted to growing-season live biomass each year. $R(T^+)$ is a randomly generated resource pulse from a log-normal distribution with mean $R_\mu$ and variance $R_{\sigma^2}$ (same mean and variance used to generate daily pulses). Our formulation assumes that at the end of each season all accumulated living biomass [$N(T^-)$] is converted to dormant biomass. Parameter notations and definitions are in table 1.

@@

```{r model_types, cache=FALSE, warning=FALSE, message=FALSE, eval=TRUE, fig.height=4.5, fig.width=8, fig.cap="Resource uptake functions and example time series of (un)correlated germination fractions for the storage effect (*a,b*) and relative nonlinearity (*c,d*) formulations of the consumer-resource model. The resource uptake functions for both species are equivalent for the storage effect, but their germination fractions are uncorrelated in time. The opposite is true for relative nonlinearity: the two species have unique resource uptake functions, but their germination fractions are perfectly correlated in time."}
library(ggplot2)
library(ggthemes)
library(mvtnorm)
library(gridExtra)
source("../modelCode/semi_discrete_consumer_resource_fxns.R")

## STORAGE EFFECT PLOTS
parms <- list(
  r = c(5,5),         # max growth rate for each species
  alpha = c(5,5),     # rate parameter for Hill function 
  beta = c(20,20),    # shape parameter for Hill function
  mN = c(0.5,0.5),    # live biomass loss (mortality) rates 
  mD = c(0.01, 0.01), # dormant biomass loss (mortality) rates
  a = c(0.5,0.5)      # allocation fraction of live biomass to seed bank
)
R <- seq(0,100,1)
out_r <- matrix(ncol=2, nrow=length(R))
alpha <- parms$alpha
beta <- parms$beta
for(i in 1:nrow(out_r)){
  out_r[i,1] <- uptake_R(parms$r[1], R[i], alpha[1], beta[1])
  out_r[i,2] <- uptake_R(parms$r[2], R[i], alpha[2], beta[2])
}
strg_uptake <- data.frame(species=rep(c(1:2), each=nrow(out_r)),
                          resource=rep(R, times=2),
                          uptake=c(out_r[,1], out_r[,2]))
strg_uptake_plot <- ggplot()+
  geom_line(data=strg_uptake, aes(x=resource, y=uptake, color=as.factor(species), size=as.factor(species), linetype=as.factor(species)))+
  theme_few()+
  scale_color_manual(values=c("#9D6188","#97A861"), name="Species")+
  scale_size_manual(values=c(3,1))+
  guides(size=FALSE, linetype=FALSE)+
  xlab("Resource")+
  ylab("Resource Uptake Rate")+
  theme(legend.position=c(0.7,0.35))+
  geom_text(aes(x=1,y=5,label="a"))

sigE <- 2   # environmental cue variability
rho <- 0    # environmental cue correlation between species
gVec <- getG(sigE = sigE, rho = rho, nTime = 50)
germ_rate <- data.frame(species=rep(c(1:2),each=50),
                        timestep=rep(c(1:50), times=2),
                        germrate=c(gVec[,1], gVec[,2]))
strg_germ_plot <- ggplot()+
  geom_line(data=germ_rate,size=1,aes(x=timestep, y=germrate, color=as.factor(species), linetype=as.factor(species)))+
  theme_few()+
  scale_color_manual(values=c("#9D6188","#97A861"), name="Species")+
  guides(size=FALSE, linetype=FALSE, color=FALSE)+
  xlab("Simulation Year")+
  ylab("Germination Fraction")+
  geom_text(aes(x=1,y=1,label="b"))

## RELATIVE NONLINEARITY PLOTS
parms <- list(
  r = c(5,1),          # max growth rate for each species
  alpha = c(5,2),      # rate parameter for Hill function 
  beta = c(20,2.5),    # shape parameter for Hill function
  mN = c(0.5,0.5),     # live biomass loss (mortality) rates 
  mD = c(0.01, 0.01),  # dormant biomass loss (mortality) rates
  a = c(0.5,0.5)       # allocation fraction of live biomass to seed bank
)
R <- seq(0,100,1)
out_r <- matrix(ncol=2, nrow=length(R))
alpha <- parms$alpha
beta <- parms$beta
for(i in 1:nrow(out_r)){
  out_r[i,1] <- uptake_R(parms$r[1], R[i], alpha[1], beta[1])
  out_r[i,2] <- uptake_R(parms$r[2], R[i], alpha[2], beta[2])
}
reln_uptake <- data.frame(species=rep(c(1:2), each=nrow(out_r)),
                          resource=rep(R, times=2),
                          uptake=c(out_r[,1], out_r[,2]))
reln_uptake_plot <- ggplot()+
  geom_line(data=reln_uptake, aes(x=resource, y=uptake, color=as.factor(species), size=as.factor(species), linetype=as.factor(species)))+
  theme_few()+
  scale_color_manual(values=c("#9D6188","#97A861"), name="Species")+
  scale_size_manual(values=c(1,1))+
  guides(size=FALSE, linetype=FALSE, color=FALSE)+
  xlab("Resource")+
  ylab("Resource Uptake Rate")+
  geom_text(aes(x=1,y=5,label="c"))

sigE <- 2   # environmental cue variability
rho <- 1    # environmental cue correlation between species
gVec <- getG(sigE = sigE, rho = rho, nTime = 50)
germ_rate <- data.frame(species=rep(c(1:2),each=50),
                        timestep=rep(c(1:50), times=2),
                        germrate=c(gVec[,1], gVec[,2]))
reln_germ_plot <- ggplot()+
  geom_line(data=germ_rate,size=1,aes(x=timestep, y=germrate, color=as.factor(species), linetype=as.factor(species), size=as.factor(species)))+
  theme_few()+
  scale_color_manual(values=c("#9D6188","#97A861"), name="Species")+
  scale_size_manual(values=c(4,1))+
  guides(size=FALSE, linetype=FALSE, color=FALSE)+
  xlab("Simulation Year")+
  ylab("Germination Fraction")+
  geom_text(aes(x=1,y=1,label="d"))

g_strg <- grid.arrange(strg_uptake_plot, strg_germ_plot, 
                       reln_uptake_plot, reln_germ_plot,
                       ncol=2, nrow=2)
```

### Model Simulations
For each coexistence mechanism (storage effect, relative nonlinearity), we ran model simulations at different levels of resource variance ($R_{\sigma^2}$), environmental cue variance ($\sigma_E$), and correlation of species responses to the environmental cue ($\rho$). 
We ran simulations for 2,000 'years' with 60 day growing seasons. 
We averaged biomass over the growing season.
After discarding an initial 100 years to reduce transient effects on our results, we calculated the stability of summed species biomass through time and species synchrony through time.
For storage effect simulations, species were equivalent in their ability to capture and assimilate the resource (*R*) but had different, depending on $\rho$, germination rates.
For relative nonlinearity simulations, species had equivalent germination rates ($\rho = 1$) but different resource acquisition functions (fig. 1).
We explored a limited set of parameter space where the storage effect and relative nonlinearity were operating simultaneously.
Within-season dynamics were solved given initial conditions using the package 'deSolve' [@Soetaert2010] in R [@Team2013].
